### 2019-10-02 docs/ICTT2019/slides/backup_slides/otherdiffdefs.tex

\begin{frame}
  \frametitle{Other diffusive-like approximations from radiative transfer problems}
  \begin{block}{Flux-limited diffusion}
  {
  The angular flux is factored in a term slowly varying in space and a normalized angular component \cite{levermore1981flux}:
  \[
  \varphi \approx \phi(\mathbf{r}, E) \psi(\mathbf{r}, \bOm) \quad \text{with} \quad \int_{4\pi} {d \bOm \psi} = 1,
  \quad \implies \mathbf{J}_g = \mathbf{f}(\mathbf{r})\,\phi_g \; \wedge \;
  \frac{\lVert \mathbf{J}_g \rVert}{\phi_g} = \text{const},\; \forall g.
  \]
  The diffusion coefficient (``the flux-limiting parameter'') come from the solution of a \emph{transcendental equation}. In the limit of weak gradients, the diffusion coefficient is obtained as (for all groups):
  \[ \textcolor{ceared1}{
  D_g(\mathbf{r}) = \frac{1}{3} \left[ \Sigma_{t,g} - \sum_{g'}{\Sigma_{s,1,g' \rightarrow g} \,\frac{\phi_{g'}}{\phi_g}} \right]^{-1}}
  \]
  }
  \end{block}
  %
  \begin{block}{Variable Eddington factor}
  {
  Postulate a factorization of the second moment by another factor (or \emph{tensor}) slowly varying in space \cite{pomraning1982flux}:
  \[
  \varphi \approx E(\mathbf{r}[, \ast]) \phi(\mathbf{r}[, \ast]),
  \text{ and }
  \partial_{\mathbf{r}} E \approx 0 \hfill \implies
  \textcolor{ceared1}{
  D_{i,g}(\mathbf{r}) = E_g(\mathbf{r}) \left[ \Sigma_{t,g} - \sum_{g'}{\Sigma_{s,1,g' \rightarrow g} \,\frac{J_{i,g'}}{J_{i,g}}} \right]^{-1}, i \in (\hat{e}_x, \hat{e}_y, \hat{e}_z).}
  \]
  }
  \end{block}
\end{frame}