### 2019-10-02 docs/ICTT2019/slides/backup_slides/slabex_Tomatis2011.tex

\begin{frame}[allowframebreaks]
  \frametitle{An example in the bare slab}
  {\hfill \cite{tomatis2011application}}%\\[3mm]

  \begin{columns}
  \column{.5\textwidth}
  \begin{block}{}
  In 1-D, we find the \textcolor{ceablue1}{exponential integral functions $E_n$} and the diffusion tensor simplifies in a single coefficient.
  \[
    \begin{split}
    E_n(\mathcal{L}_{op}) = \int_1^{\infty}{du\, \frac{e^{-|\mathcal{L}_{op}|\,u}}{u^n}}\\
      = \int_0^1 {d\mu \, \mu^{n-2} e^{-|\mathcal{L}_{op}| / \mu}}.
    \end{split}
  \]
  {\textcolor{ceablue1}{APPROX}: 2G problem with isotropic scattering, extrapolated length $\zeta = 2.13 \mathcal{D}$}\\
  {\textcolor{ceablue1}{b.c.}: reflection at left, vacuum or zero flux at right.}
  \end{block}
  \column{.5\textwidth}
  \begin{table}\centering
  \caption{17$\times$17 PWR UO2 FA data (3.5\% enr. w/o $^{235}$U).}
  \begin{tabular}{lc} \hline \hline
    Slab width & 21.5 cm\\
    Fuel temp. & 900 K\\
    Water temp.& 600 K\\
    Water density& 0.66 g/cm$^3$\\
    Boron conc.& 500 ppm\\ \hline \hline
  \end{tabular}
  \end{table}
  \end{columns}

  \begin{block}{Results for the $k$-eigenvalue and the flux convergence}
  \centering
  \begin{tabular}{*{4}{c}} \hline \hline
   Left b.c.: & \multicolumn{2}{c}{vacuum} & zero flux\\
   \cline{2-4}
    & $k$ & $\Delta k$ (pcm) & $\Delta k$ (pcm) \\ \hline
    Ref. S16             & 0.745675 & ~~~0~ &    -    \\
    $D = 1/(3 \Sigma_t)$ & 0.741377 & -429.8& -5800.1 \\
    $D = \Sigma_s/(3 \Sigma_t^2)$ & 0.746360 & ~~68.5& -5150.1 \\
    Ronen corr. & 0.743860 & \textcolor{ceablue1}{-181.5} & \textcolor{ceablue1}{164.4}\\
    \hline \hline
  \end{tabular}
  \end{block}

  \begin{figure}
    \includegraphics[width=.7\textwidth]{figures/Tflx_cnv.png}
    \caption{Thermal flux}
  \end{figure}

  \begin{figure}
    \includegraphics[width=.7\textwidth]{figures/Fflx_cnv.png}
    \caption{Fast flux}
  \end{figure}

  \begin{table}
  \caption{Cross section data of the homogenized FA ($k_{\infty} = 1.07838$); scattering cross section comes from sum on columns.}
  \begin{tabular}{*{6}{c}}\hline \hline
  Group $g$ & $\Sigma_{t,g}$ & \multicolumn{2}{c}{$\Sigma_{s,0,g' \rightarrow g}$} &
  $\chi_g$ & $\nu\Sigma_f$ \\ \hline
  1 & 5.31150E-01 & 5.04664E-01 & 2.03884E-03 & 1 & 7.15848E-03 \\
  2 & 1.30058E+00 & 1.62955E-02 & 1.19134E+00 & 0 & 1.41284E-01 \\
  \hline \hline
  \end{tabular}
  \end{table}

  \begin{exampleblock}{Discussion}\tt
  \begin{itemize}
    \item The Ronen method drives the diffusion solution towards a transport-like solution even with unphysical b.c.
    \item The error concentrates at the boundary interface, and asks for many iterations
    \item $D=1/(3\Sigma_t)$ is the $P_1$ diffusion coefficient here
    \item Although $D=\Sigma_s/(3\Sigma_t^2)$ results in a smaller $\Delta k$, $\Delta \phi$ is 20\% off near the left boundary
  \end{itemize}
  \end{exampleblock}

  \begin{alertblock}{Same issues and remarks for multi-dimensional $n$-D cases}\tt
  \begin{itemize}
    \item Can we prove convergence?
    \item Can we compute all components of the diffusion tensor in a coarse mesh?
    \item What are the special functions in $n$-D?
    \item Are there regions (like \emph{void} for instance) where the method cannot be applied?
  \end{itemize}
  \end{alertblock}
\end{frame}