### 2019-10-02 docs/ICTT2019/slides/backup_slides/sourcebymoments.tex

\begin{frame}
  \frametitle{Treatment of scattering anisotropy as in ENDF}
  \begin{block}{Introducing the flux moments on spherical harmonics in the NTE}
  {
  \begin{gather*}
  \left[ \frac{1}{v}\frac{\partial}{\partial t}
  %+ \nabla \cdot (\bOm \,\_) + \Sigma_t(\mathbf{r}, E) \right] \varphi(\mathbf{r}, E, \bOm, t)
  + \bOm \cdot \nabla  + \Sigma_t(\mathbf{r}, E) \right] \varphi(\mathbf{r}, E, \bOm, t)
  = \frac{\chi(E)}{4\pi} \int_0^{\infty} {dE' \nu\Sigma_f(\mathbf{r}, E') \phi(\mathbf{r}, E, t')}
  + S_{ext}(\mathbf{r}, E, \bOm, t) +\\
  \visible<2-3> {\frac{1}{2\pi}}
  \only<-3>{\int_{4\pi} d\bOm'}
  \only<4->{ \sum_{\ell = 0}^{\infty} \sum_{m=-\ell}^{\ell}  Y_{\ell m}(\bOm) }
  \int_0^{\infty} {dE' \only<-3>{\Sigma_s}\only<4->{\Sigma_{s, \ell}}
  (\mathbf{r}, E' \rightarrow E
  \only<1> {, \bOm' \rightarrow \bOm)}
  \only<2-3>{, \textcolor{ceablue1}{\bOm' \cdot \bOm})}
  \only<-3>{\varphi(\mathbf{r}, E', \bOm', t)}
  \only<4->{)\phi_{\ell m}(\mathbf{r}, E', t) }
  },
  \end{gather*}
  }
  {\sl hp.}: a unique family of fissile nuclides, all prompts -- b.c. at $\mathbf{r}_S$: $\varphi(\mathbf{r}_S, E, \bOm) = 0$ if $\hat{n} \cdot \bOm < 0$, $\forall E$. \uncover<2->{$\mu_0 = \bOm \cdot \bOm'$ is the cosine of the scattering angle.} \uncover<3->{Thanks to the addition theorem $P_0$ is written in terms of the spherical harmonics functions $Y_{\ell m}$}
  \uncover<2->{ \[
    \Sigma_s(\mathbf{r}, E' \rightarrow E, \mu_0) = \sum_{\ell = 0}^{\infty}{\frac{2\ell + 1}{2}\Sigma_{s,\ell}(\mathbf{r}, E' \rightarrow E) P_\ell(\mu_0)}
   \visible<3-> {= 2\pi \sum_{\ell = 0}^{\infty} {\sum_{m=-\ell}^{\ell}{ \Sigma_{s,\ell} Y^*_{\ell m}(\bOm') Y_{\ell m}(\bOm)}}}
  \]}
  \uncover<3->{\emph{Spherical harmonics} in $\bOm = (\mu, \theta)$, defined by associated Legendre polynomials $P_{\ell}^m$:
  \[ Y_{\ell m}(\bOm) = \left[ \frac{2 \ell+1}{4\pi} \frac{(\ell - |m|)!}{(\ell + |m|)!} \right]^{1/2} P_{\ell}^m (\mu)e^{i m \theta}
  \uncover<4->{ \quad
  \text{so that } \textcolor{ceablue1}{
  \varphi = \sum_{\ell = 0}^{\infty} \sum_{m=-\ell}^{\ell} \left( \frac{2\ell + 1}{4\pi} \right)^{1/2} \phi_{\ell m}(\mathbf{r}[, \ast]) Y_{\ell m}(\bOm)}}
  \]
  and subject to the orthogonal-normalization property $\int_{4\pi} {d \bOm Y^*_{\ell m}(\bOm) Y_{k n}(\bOm)} = \delta_{\ell k}\delta_{m n}$.}
  \end{block}
\end{frame}