### 2020-06-05 docs/PHYSOR2020/ANE-RM-2020.tex

\documentclass[preprint, authoryear, 12pt, a4paper, onecolumn]{elsarticle}
%
%  various packages that you may wish to activate for usage 
%\usepackage{booktabs}
%\usepackage{natbib}
%\usepackage{cite}
%\usepackage{epsf}
\usepackage{bm}
\usepackage[english]{babel}
%\usepackage[autostyle, english = american]{csquotes}
%\MakeOuterQuote{"}
\usepackage{color}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{calrsfs}
\usepackage{lineno}
\usepackage[colorlinks=true, a4paper=true, pdfstartview=FitV,%
            linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}

% for algorithms
% \usepackage{algorithm}
% \usepackage{algpseudocode}

\usepackage{tikz}
\usepackage{xfrac}
\usepackage{units}

\usepackage[printonlyused]{acronym}

\journal{Annals of Nuclear Energy}

\newcommand{\eq}[1]{Eq.~(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.~(\ref{#1})}
% \newcommand{\tsup}[1]{\textsuperscript{#1}}
% \newcommand{\tsub}[1]{\textsuberscript{#1}}
\newcommand{\keff}{\ensuremath{k_{\textrm{\scriptsize{eff}}}}}
\newcommand{\kinf}{\ensuremath{k_{\textrm{\scriptsize{$\infty$}}}}}
\newcommand{\pf}{\textrm{PPF}}
\newcommand{\eg}{e.g.,}
\newcommand{\ie}{i.e.,}
\newcommand{\OO}{\hat{\bm{\Omega}}}
\newcommand{\rr}{\bm{r}}
\newcommand{\jp}{\ensuremath{J^+}}
\newcommand{\jm}{\ensuremath{J^-}}
\newcommand{\jpm}{\ensuremath{J^\pm}}
\newcommand{\delj}{\ensuremath{\delta J}}
\newcommand{\jD}{\ensuremath{J^{\textrm{\scriptsize{D}}}}}
\newcommand{\bmj}{\ensuremath{\bm{J}}}
\newcommand{\rE}{\ensuremath{(\rr,E)}}
\newcommand{\bnabla}{\ensuremath{\bm{\nabla}}}
\newcommand{\hzi}{\ensuremath{\sfrac{1}{2}}}
\newcommand{\mcL}{\tau}
\newcommand{\dx}{\Delta}
\newcommand{\dd}{\delta D}
\newcommand{\tild}{\tilde{D}}

% about notes and comments from each author
\newcommand{\erez}[1]{{\color{blue}{\textsuperscript{EREZ:}#1}}}
\newcommand{\roy}[1]{{\color{blue}{\textsuperscript{ROY:}#1}}}
\newcommand{\daniele}[1]{{\color{blue}{\textsuperscript{DANIELE:}#1}}}
\newcommand{\revised}[1]{{\color{red}{#1}}}

\newcommand\mscriptsize[1]{\mbox{\scriptsize\ensuremath{#1}}}
\newcommand\mtiny[1]{\mbox{\tiny\ensuremath{#1}}}

\renewcommand{\cite}{\citep}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   BEGIN DOCUMENT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{frontmatter}

  \title{Integral Transport Correction to Diffusion Calculations \\ Using the pCMFD Scheme}

  %% or include affiliations in footnotes:
  \author[bgu]{R.~Gross}
  \author[cea]{D.~Tomatis}
  \author[bgu]{E.~Gilad\corref{cor1}}
  \ead{gilade@bgu.ac.il}

  \cortext[cor1]{Corresponding author}

  \address[bgu]{The Unit of Nuclear Engineering, Ben-Gurion University of the Negev, Beer-Sheva 8410501, Israel}
  \address[cea]{CEA, DEN, Service d{'}\'etudes des r\'eacteurs et de math\'ematiques appliqu\'ees (SERMA), \\
                Universit\'e Paris-Saclay, F-91191, Gif-sur-Yvette, France}

  %------------------------------------------------------------
  \input{abstract/abstract}
  %------------------------------------------------------------
  \begin{keyword}
  \input{keywords}
  \end{keyword}
  %------------------------------------------------------------
\end{frontmatter}

%\clearpage

\modulolinenumbers[5]
\linenumbers

\input{contents}

% ------------------------------------------
%\section*{ACKNOWLEDGEMENTS}
%
%Acknowledge the help of colleagues and sources of funding, as appropriate.
%
% ------------------------------------------

\bibliographystyle{elsarticle-harv}
\bibliography{RM-bib}

% ------------------------------------------
%\appendix
%\gdef\thesection{APPENDIX \Alph{section}}
%\section{Sample Appendix 1}
%\label{app:a}
%If necessary, include Appendices numbered in upper case alphabetical order. This is \ref{app:a}.
%
\end{document}