### 2020-09-04 docs/ICTT2019/review/response2reviewers.txt
start revision of JCTT paper (special issue after ICTT26)

Dear Editor,

The authors are very thankful to the reviewers for their constructive work.
Please find hereafter the authors' response to the reviewers. Authors' replies follow after the keyword "As:".
The revised copy of the article shows all changes by colored text to support the reviewing process.
Best regards,

Daniele Tomatis
(on behalf of all authors)
Saclay, France
August 30th, 2020
***

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author
There are two line numberings in the article. The comments here refer to the numbering at the far left (from 1 to 60).
As: remark from the authors - the far left numbering provided authomatically by the online system of the journal seems mismatching the position of the lines in the pages. Hence, we provided our own line numbers.

Comment 1
Page 1, line 21. I think that « to solve the neutron transport equation » is better than « to resolve the neutron transport ».
As: ***.

Comment 2
Page 5, line 15. The relation giving the optical length τ is important and deserves a better presentation. Please put this relation into a separate line.
As: Changed as recommended by the reviewer.

Comment 3
Page 5, line 35. At this point it would be very useful for the reader to have the expression of q_g with the sum limited to l=0 and 1. This will allow to better understand the equations that follow.
As: The repetion of eq. 5 with the sum over $l$ truncated to 1 would take additional space with limited amount of new information, which the author would like to avoid. The sentence "Therefore, the sum over $l$ will always be truncated in the following." has been added at the end of the paragraph in order to make the text clearer.

Comment 4
Page 6, line 13. I suggest to replace:
 « The optical lengths … »
with:
« The optical lengths τ in Eq. 6 … ».
As: Suggestion accepted.

Comment 5
Page 6, line 32. Please provide the expression defining e tilde.
As: ***.

Comment 6
Page 6, line 33. Please provide the expression defining t tilde.
As: ***.

Comment 7
Page 6, line 48. Does the term « distance » mean « travel path »?
As: here distance stands for linear extent in space along the given direction between the point at r and the outer surface of a finite body. It means travel path only for neutrons starting their first flight after emission at the specific distance S'.

Comment 8
Page 7, line 11. After the sentence « The theory of collision probability methods » put the acronym (CPM). It is already given in the abstract, but in scientific articles the main text is independent of the abstract.
As: thanks for your remark. The acronym has been inserted in the text as suggested.

Comment 9
Page 7, line 51. Does the expression {ω | …} mean projection on ω? Is this a conventional formalism? Is this equivalent to the convention <… | ….> ?
As: the symbol | stands for 'such that' from set notation, see https://en.wikipedia.org/wiki/List_of_mathematical_symbols.

Comment 10
Page 9, line 10. After « h-integration » please put « in Eq. 16 ».
As: the integration on h is introduced in Eq. 13 and it is used also for the following equations. Please, see the new form of the text that uses 'introduced in Eq. 13...'.

Comment 11
Page 10, line 13. I suggest replacing « to consider » with « to take into account ».
As: replaced with 'taking into account'.

Comment 12
Page 11, line 47. I suggest replacing « reproduced » with « assumed ».
As: text replaced as suggested by the reviewer.

Comment 13
Page 12, line 38. I suggest replacing « can not » with « cannot ».
As: text replaced as suggested by the reviewer.

Comment 14
Page 13, line 30. I suggest replacing « follows by Sood et al. » with « follows Sood et al. [13]».
As: ***.

Comment 15
Page 13, line 55. I suggest replacing « column, see Table II » with « column of Table II ».
As: The sentence has been revised to get a simpler form.

Comment 16
Page 14, line 5. I suggest replacing « communicate » with « provide ».
As: suggestion accepted.

Comment 17
Page 14, line 47. After « slab thickness » put the symbol « (r_c) ».
As: text modified as requested.

Comment 18
Page 15. Fig. 5 is very difficult to understand. 
In the caption it is written: « reflection at the right and vacuum at the left ». Usually a half slab is cut at the left, so the reflexion is expected at left.
As: New elements have been added to the text to understand better the figure, see the revised copy please. We confirm reflection at the right and vacuum at left, as written in the submitted text and performed in our calculations.

- Left figure 15. L_c is never defined. Is it r_c? Moreover, when printed in black and white, it is difficult to assign the CP or RM origin to the two clusters of results. I suggest to indicate with an arrow with RM the green cluster and with CP the orange cluster.
As: L_c --> r_c; marker '+' added to identify the cluster about RM results.

- Right figure 15. The symbol Delta in the abscissa is never defined so the right figure is not easy to understand. Moreover the units of Delta should be reported. I suppose HS-CP means half-slab computed with collision probabilities. What method is used for the FS (full-slab)? I suppose RM. If so, put (FS-RM).
As: Figure completed as requested. New sentence added: 'Equidistant meshes are used by refining gradually the cell width $\Delta$.'.

Comment 19
Page 15, line 55. Please provide a reference for the Thomas TDMA algorithm.
As: citation added as requested.

Comment 20
The RM is intended to improve the diffusion theory. In order to evaluate how much it is improved, please add inside Figs. 5, 6 and 7 the results of diffusion theory: the analytical result from the equation in table 3 and the numerical result, which is the first iteration of RM. The gain in computation time with respect to CP should be done, as well as the loss with respect to pure diffusion theory.
As: The RM was initially proposed to improve diffusion theory without a formal connection to transport. In this article it is presented as a technique to solve neutron transport by means of a diffusion equation **relying on the theory of the collision probability method**. Results from diffusion theory are far from the ones obtained by transport theory, and we only report the analytical solutions for each test case, thus avoiding also the repetition of information which is already known. The evolution of the flux solution throughout the RM iterations has already been presented by Tomatis and Dall'Osso 2011 and Gross et al. 2020.

Comment 21
Page 17, line 48. Please provide a reference to support the sentence « previous published works ».
As: references added as requested.

Comment 22
About the equation numbering inside Appendices. Conventionally the equation numbering inside Appendices does not follow the numbering inside the main text of the article. So Eq. 21 should be Eq. A.1, Eq. 22 should be Eq. A.2, Eq. 24a should be Eq. B.1a, and so on.
As: This article was prepared for the ICTT26 conference using the specific template of the same conference. Some articles presented at the conference were proposed for publication on JCTT, so we continued using the same template for the final paper. The template of the journal will be used for the revised copy.

Comment 23
Page 19, line 5. It is said: « the values from these libraries are different ». Please report approximately the magnitude order of the difference.
As: Authors indicated the numerical functions and the code versions of the scipy and numpy libraries used to calculate the values. The specific commands are also available in the program published online. The advert reader can run the same commands in the python interpreter and get the sought values. Therefore, we would like to avoid printing these roots and weights values. These roots and weights are totally different from the numbers printed in the cited textbooks in reason of the different procedure used to solve the integral by numerical quadrature. It is not relevant to quantify the differences as far as the quadrature is correctly derived and computed (verification test available online at https://github.com/ndarmage/RM1D/blob/public/CPM1D/tests/CP_empty_cylinder.py).

Comment 24
Page 20, line 14. « algebraic system equations » => « algebraic system of equations ».
As: changed as indicated.

Comment 25
Page 20, line 24. For clarity, please add after « are constructed by blocks. » the sentence:
« I is the identity matrix. ».
As: changed as indicated.

Comment 26
Page 20, line 24. I suggest to replace « with all » with « and is constituted of ».
As: changed as indicated.

Comment 27
The format of the article does not match the format required by the journal. The required format can be found at:
https://authorservices.taylorandfrancis.com/tf_quick_guide/
Please verify whether the publisher accepts other formats or it is up to the author to modify the format of his article.
As: We used the template of the conference as mentioned before. Thanks for indicating the url above by which we arrived to the webpage:
https://authorservices.taylorandfrancis.com/formatting-and-templates/
Unfortunately, there are no latex template available at this page. Hence, we contacted the editor to ask for the proper template.
We are going to submit the revised copy with the proper latex template of the journal.

As: The author are very thankful to Reviwer 1 for her/his thorough work.

Reviewer: 2

Comments to the Author
Please see the attached file.

Please find the authors' response below according to the reviewer's numbering.
i) *** do you think that the title change is suitable? ***
ii) Although the numerical tests were performed only for isotropic sources, we allow for linearly anisotropy sources in the derivation of our equations. This limitation is reminded in the conclusion, where authors have added a few changes in the text to enlight the fact that more work will ne needed in the future to study appropriately source anisotropy.
iii) the derivation of the transmission probabilities and the formulae of reciprocity (section 3.4) for entering neutrons at the boundary requires white reflection, with isotropic flux (see the end of the paragraph after eqs. 19).
iv) done

Editor's Comments to Author:

Associate Editor
Comments to the Author:
(There are no comments.)